\section{Failure Modes}

\begin{lemma}[Late-response failure]
If the agent waits until inertia has fully rebuilt after a window closes, the
same target action requires strictly greater effort than during the window.
\end{lemma}

\begin{corollary}[Verbal commitment is mechanically weak without action]
Endorsing a future action in language has limited strategic value if no
stabilizing movement begins before the decision window expires.
\end{corollary}

\begin{proposition}[Preparation failure]
Even when a decision window opens, the transition may fail if missing objects,
unclear next steps, or unresolved setup costs consume the interval before the
target state begins to stabilize.
\end{proposition}

\begin{proposition}[Restabilization failure]
Initial movement toward a new state is not sufficient for a successful
transition. If the new regime does not receive early stabilizing action, the
previous basin can quickly regain control.
\end{proposition}

\begin{discussion}
This lemma explains a common misunderstanding. People often believe they have
``decided'' to act because they endorsed a future action verbally. But if they
do not move inside the decision window, the state soon becomes expensive again
and the earlier verbal decision has almost no mechanical value.
\end{discussion}

\begin{example}[The document opens but the regime does not change]
A reader opens the paper they planned to study, but the first concrete task is
not preselected. Tabs, notifications, and uncertainty refill the environment
before the new task state can stabilize. From the outside this looks like weak
discipline; mechanically it is a failure to convert initial movement into a
self-supporting regime.
\end{example}

\begin{remark}[Recovery rule]
After a missed window, the highest-value response is usually not to renegotiate
identity or moral worth. The better move is to prepare the next window by
removing one upstream barrier variable: pre-open the document, pack the bag,
write the first line, or schedule the social trigger that makes the next entry
cheaper.
\end{remark}
