\section{Recommended Reading}

\begin{readingbox}
\textbf{Foundation}
\begin{itemize}[nosep]
  \item Eugene Izhikevich, \emph{Dynamical Systems in Neuroscience}, for
        disciplined state-space and stability language.
  \item Michael Brin and Garrett Stuck, \emph{Introduction to Dynamical
        Systems}, for mathematical background on state evolution and local
        structure.
  \item Steven Strogatz, \emph{Nonlinear Dynamics and Chaos}, for threshold,
        stability, and regime-change intuition without overclaiming empirical
        fit.
\end{itemize}
\textbf{Modern Research and Review}
\begin{itemize}[nosep]
  \item Review literature on attractor and integrator networks in the brain
        (Nature Reviews Neuroscience, 2022, medium), for why state persistence
        and retained directional information are serious modelling ideas.
  \item \emph{Metastability demystified: the foundational past, the pragmatic
        present and the promising future} (Nature Reviews Neuroscience, 2024,
        medium), for why stable-but-revisable state descriptions are often more
        appropriate than crude rigid-versus-fluid binaries.
  \item \emph{Cortical connectivity, local dynamics and stability correlates of
        global conscious states} (Communications Biology, 2025, strong as a
        biological legitimacy source, not as direct validation of this
        chapter's scoring formula).
\end{itemize}
\textbf{Reading discipline}
\begin{itemize}[nosep]
  \item Use the textbooks to justify the modelling language.
  \item Use the review and biological sources to justify taking state,
        stability, and transition structure seriously.
  \item Do not read either group of sources as direct proof that this chapter's
        heuristic score has already been empirically calibrated.
  \item If later chapters introduce stronger formulas, they must explicitly say
        whether those formulas remain heuristic, merely monotone, or genuinely
        supported beyond what is established here.
\end{itemize}
\end{readingbox}
