\section{Case 2: Exercise Non-Start After Work}

\subsection*{Situation}
An employee endorses regular exercise every morning but rarely reaches the gym
after a draining workday.

\subsection*{Framework map}
\begin{itemize}[nosep]
  \item \textbf{Current state:} end-of-work depletion plus home-directed drift.
  \item \textbf{Target state:} arrival at the gym and completion of a prepared
        session.
  \item \textbf{Dominant variables:} low $\Will$, elevated $\DV$, and
        environment-dependent re-locking into a rest regime.
  \item \textbf{Missed decision window:} commute boundary before home entry.
\end{itemize}

\subsection*{Mechanism interpretation}
This case illustrates how timing and path dependence compound one another. If
the commute reliably terminates in the home-rest sequence, the trajectory is
already partially selected before conscious deliberation begins. In the
language of advanced dynamics, the home route behaves like a basin-reentry
channel. In the language of emergence, the repeated work-to-home-to-collapse
sequence becomes a historically consolidated transition policy.

\subsection*{Evidence lane discipline}
The current research spine supports a \textbf{medium} interpretation that
stability architecture and transition timing matter. It does \emph{not}
support a strong claim that evening exercise failure has been mapped onto a
specific neural attractor with clean empirical identification. The biologically
grounded lesson is narrower: state persistence, stability, and transition cost
are serious system properties, so intervention should target those properties
first.

\subsection*{Operational repair}
\begin{enumerate}[nosep]
  \item Move the decision upstream by packing the gym bag before work.
  \item Choose one default workout rather than a menu of options.
  \item Route the commute past the gym instead of directly home.
  \item Define success as physical entry plus one pre-specified minimum block
        of exercise, not an idealized full workout.
\end{enumerate}

\subsection*{Transfer lesson}
This case teaches a general rule: if the current route already reconstructs
the old state before deliberation starts, redesign the route rather than asking
the depleted self to overcome the full barrier in place.
