\section{Case 1: Study Start Failure}

\subsection*{Situation}
A student returns from lunch intending to read a difficult chapter but instead
opens short-video feeds for forty minutes.

\subsection*{Framework map}
\begin{itemize}[nosep]
  \item \textbf{Current state:} post-lunch passive novelty consumption.
  \item \textbf{Target state:} seated proof-focused reading with active note
        production.
  \item \textbf{Dominant variables:} high $\Reward$, high $\Switch$, and high
        $\Env$.
  \item \textbf{Missed decision window:} immediate re-entry interval after the
        meal, before the phone reconstructs the old regime.
\end{itemize}

\subsection*{Mechanism interpretation}
This case is best read as a local state-transition failure. The current state
has an easily re-entered attractor-like character: low effort, rapid novelty,
and a familiar posture-tool configuration. The target state is not absent only
because the student lacks values; it is weakly stabilized and poorly prepared.
The emergence chapter adds a second layer: repeated post-lunch phone use can
create path dependence, so the same temporal boundary starts to function as a
history-loaded cue rather than a neutral moment.

\subsection*{Evidence lane discipline}
The strong empirical consciousness sources support only a limited transfer
here: organized states and transitions can exhibit structured stability and
changing dynamics. The case therefore uses \textbf{strong-to-medium}
translation, not direct proof. The sleep-transition bifurcation paper is
relevant as a model for why transition timing matters, but it does not prove
that every lunch break is a bifurcation in the strict neuroscientific sense.
Metastability language is useful in a \textbf{medium} sense: the work state
must become stable enough to hold, yet still be easy to recover after
interruption.

\subsection*{Operational repair}
\begin{enumerate}[nosep]
  \item Write the exact next proof target before the meal break.
  \item Leave the notebook open at the unfinished line.
  \item Remove the phone from the immediate work surface.
  \item Define the first post-break action as ``sit, read the last two lines,
        and write one continuation sentence.''
\end{enumerate}

\subsection*{Transfer lesson}
Study-start failure is usually not solved by increasing moral pressure. It is
solved by making the desired re-entry trajectory shorter, more visible, and
less dependent on fresh design at the moment of return.
