\section{Case 3: Writing Avoidance in Research Work}

\subsection*{Situation}
A graduate student can read papers and discuss ideas but repeatedly delays
writing drafts.

\subsection*{Framework map}
\begin{itemize}[nosep]
  \item \textbf{Current state:} browsing, annotating, and discussing without
        visible output.
  \item \textbf{Target state:} rough written production with explicit claims.
  \item \textbf{Dominant variables:} moderate $\Reward$ in pseudo-work, very
        high $\Switch$ due to ambiguity, and self-protective environmental
        drift toward low-exposure tasks.
  \item \textbf{Missed decision window:} first minute of the writing session,
        when the draft could still be defined narrowly.
\end{itemize}

\subsection*{Mechanism interpretation}
This is a neighbouring-attractor case rather than a simple inactivity case.
The student is active, but active in a basin that provides intellectual
legitimacy without forcing externalized structure. Emergence language matters
because repeated success in reading or discussing can produce a path-dependent
identity loop: the system becomes efficient at the pre-draft regime and
therefore increasingly expensive to shift into visible production. Attractor
language helps here because the problem is not zero movement; it is movement
within the wrong region of state space.

\subsection*{Evidence lane discipline}
The mechanism claim here is \textbf{medium}. The attractor and metastability
literature supports using basin, persistence, and reconfiguration language as a
formal scaffold. It does not warrant a strong empirical claim that one can read
off ``writing avoidance'' directly from brain-level measurements. The case
therefore remains a disciplined behavioural model supported by systems
concepts, not a completed laboratory result.

\subsection*{Operational repair}
\begin{enumerate}[nosep]
  \item Redefine the target as one ugly but visible unit: five bullet claims,
        one paragraph skeleton, or one figure caption.
  \item Open the session with a production artifact, not a reading artifact.
  \item Keep one unresolved question from the previous session visible.
  \item Separate evaluation from generation so that sentence production does
        not have to carry full quality judgment at entry.
\end{enumerate}

\subsection*{Transfer lesson}
Some failures arise because the system has become too competent at an adjacent
state. In such cases the intervention must reduce target ambiguity and expose
the smallest real output unit, rather than merely extending time spent near the
target domain.
